### 2021-06-06 Rmd/count_check_MS.Rmd
improved methods, bird data

---
title: "Tools for identifying unexpectedly low microfossil count sums"
author:
  - name: Richard J. Telford
    email: richard.telford@uib.no
    affiliation: University of Bergen
    footnote: Corresponding Author
address:
  - code: University of Bergen
    address: Department of Biological Sciences, University of Bergen and Bjerknes Centre for Climate Research, Post Box 7803, N-5020 Bergen, Norway
abstract: |
  Microfossil counts are a key data type in palaeoecology. Recent work has raised the possibility that some authors might misreport an important quality control parameter, the counts sums, occasionally dramatically so, and perhaps other aspects of the counts. This paper explores the statistical properties of ecological and palaeoecological species counts and uses the findings to develop methods to flag count or percentage data with unexpected distributions. Several assemblage datasets that fail these tests, and potentially have misreported counts or count sums, are identified.
journal: "Journal of Quaternary Science"
date: "`r format(Sys.Date(), '%d %B %Y')`"
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}   
output:
  bookdown::pdf_book:
    base_format: 
      rticles::elsevier_article
    number_sections: true
documentclass: article
classoption: a4paper
bibliography: extra/countMS2.bib
csl: extra/elsevier-harvard_rjt.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, warning = TRUE, message = TRUE, error = TRUE)

library(drake)
library(tidyverse)
library(english)
library(rjt.misc)

#set ggplot theme
theme_set(theme_bw(base_size = 12))
options(digits = 2)

loadd(bird_singleton_summary, bird_summ)
loadd(testate_summ)
loadd(last_chance_summ)
loadd(pollen_summ, hershop_pollen_summ, wildhorse, pollen_summ1, pollen_summ2, pollen3, pollen1, pollen2)
loadd(owens_summ1, owens_summ, owens_est_summ)
loadd(diatom1_summ, diatom2_summ, chironomid1_summ, marine1_summ, chironomid2_summ)

```

**Keywords:** Microfossils assemblages, Counts; Quantitative methods, Reproducible research, Meta-research

# Introduction

Six percent of papers published in _Molecular and Cellular Biology_ show evidence of inappropriate image duplication [@Bik2018] either due to error or, more rarely, misconduct.
It would be reckless to assume that the palaeoecological literature does not have equivalent problems: the low rate of retractions in the ecological and geological literature [@Grieneisen2012] may partially reflect our limited ability to detect errors and misconduct rather than their prevalence.

Several numerical tools have been developed to identify questionable data.
Deviations from the distribution of digits expected from the Newcomb-Benford law has been used to detect issues with scientific and financial data [@Barabesi2017].
@Carlisle2017 identified papers where the baseline differences in means for different treatments were surprisingly high or low given the variance of the data.
@Brown2017 developed the granularity-related inconsistency of means (GRIM) test which assesses whether means of integer data are consistent with the reported sample size.
However none of these tests are directly applicable to microfossil assemblage data, one of the most common types of palaeoecological data.

A common assertion in papers reporting microfossil assemblage data is that a minimum of _N_ microfossils were counted, where _N_ is often fifty for chironomids and several hundred for pollen and diatoms.
This is important because larger count sums are associated with smaller uncertainties, both in relative abundance of taxa and derived statistics such as transfer function reconstructions [@Heiri2001; @Payne2009].
There is a risk that count sums will be misreported.
This could happen by mistake, especially with the use of secondary data.
It could also happen knowingly, especially when microfossil preservation is poor or concentrations are low.
This risk that count sums will be misreported is demonstrated by @Larocque-Tobler2016 who acknowledged that the chironomid count sums in @Larocque-Tobler2015 were not at least fifty head capsules as reported but as low as nineteen.
@Telford2019 reports some other cases where count sums may be much lower than reported.

Given that count sums of percent data have been misreported, the possibility that other aspects of count data may be misreported needs to be considered.

The key insight that allows inference about community or assemblage count data is that they represent a species abundance distribution.
The ubiquitous shape of species abundance distributions when plotted as the number of species against abundance is a hyperbolic or hollow curve with many rare and few abundant species and is central to ecological theory [@McGill_2007].
From this, two predictions can be made: that counts of an entire community or assemblage should have many rare taxa, and that the distribution of abundances should be smooth.
Count data that differ markedly from these predictions might not accurately reflect the original community or assemblage.

This paper aims to develop and test simple methods to test the abundance of rare taxa and the smoothness of the species abundance distribution that can  use to flag cases if the count data might have been misreported.
Next, as, regrettably, many palaeoecological data sets are archived as percent data, these tests are then extended to identify percent data where the count sum might be lower than reported.
Finally, some complications and caveats are discussed.

<!-- For example, because there are many rare taxa in most communities [@Darwin1859], most community or assemblage samples will include taxa represented by a single individual, hereafter singletons, unless the sampling effort is high relative to the taxonomic richness [@Coddington2009]. -->

<!-- In the occasional assemblages without singletons, the greatest common divisor should usually be one, i.e. few assemblages should have all counts divisible by an integer larger than one, especially if species richness is high. -->

# Methods

## Data

Publicly available datasets were downloaded from the Palaeodata Center, Neotoma, Pangaea, and other sources.
A range of ecological and palaeoecological data were sought to allow for differences in typical count sums and species richness.

Count data analysed include the North American Breeding Bird survey [@bbsAssistant2]; 
the Brown University planktic foraminiferal database [@Prell1999];
the MOLTEN diatom database [@Elinor]; 
the North American testate amoeba training set [@AMESBURY2018];
and fossil pollen assemblage from the Neotoma database [@neotoma_2018]. 
Pollen data were downloaded on `r readd(pollen_download_date)`, samples with with small count sums (< 50), that appeared to be percentages, or were documented as back-transformed from digitised data were excluded.

Examples of percent data where the minimum count sum was reported in the associated paper were sought.
Datasets discussed by @Telford2019 as possibly having misreported the minimum count sum were also excluded to avoid double reporting.
Datasets with possible misreporting are anonymised; no attempt is made to diagnose whether errors or misconduct are responsible.
This paper does not attempt to be an exhaustive survey of all the data available.

## Numerical methods

### Test for count data

To test for rare taxa, the prevalence of samples without singletons was calculated for each data set.

To test if the species abundance distribution is smooth, I used two tests.
First, for each assemblage or community with at least ` word(min_richness)` taxa, the proportion of taxa with counts divisible by $k$ where $k$ is a small integer larger than one is calculated.
Samples where this proportion is high have a saw-toothed pattern of counts. 
In extreme cases, where all taxa have counts divisible by $k$, the sample has a greatest common divisor of $k$.
Data sets with a many samples with a greatest common divisor greater than one or an unusually high proportion, relative to the majority of pollen data set in the Neotoma database, of counts divisible by $k$ were flagged for closer inspection.

The second test of smoothness is to fit a generalised additive model with Poisson errors to the counts aggregated over samples in each data set. 
Only counts of ` word(gam_max)` or fewer were considered due to low power in the relatively infrequent larger counts.
Data sets where the proportion of explained deviance is low relative to the majority of pollen data set in the Neotoma database were flagged for closer inspection.

### Test for percent data

In general, it is not possible to determine the sum from which percentages were calculated, but the properties of community and assemblage counts make it possible to estimate it.
Given the expectation that the rarest taxon is a singleton, the count sum _N_ can be estimated as $1/p_{min} \times 100$ where _p~min~_ is the percent abundance of the rarest taxon.
This method will fail for assemblages without singletons.
It is also expected that the percent _p_ is calculated from integer counts, therefore there should be a count sum _N_ such that $p_i/100\times N$ is, within rounding error, an integer for all _i_ taxa.
Possible values for _N_ can be found by a direct search algorithm over the range of plausible values of _N_.
An infinite number of possible count sums that are consistent with the percent exist, but the lowest will give the correct value of _N_ except in cases where the greatest common divisor is greater than one.
These tests for percent data are closely related to GRIM [@Brown2017], as all rely on the granularity of percentages calculated from integer data.

### Software

All analyses were done in R v. `r getRversion()` [@R] and used the packages 
extraDistr v. `r packageVersion("extraDistr")` [@extraDistr],
numbers v. `r packageVersion("numbers")` [@numbers], 
countSum v. `r packageVersion("countSum")` [@countSum], 
neotoma v. `r packageVersion("neotoma")` [@neotoma], 
bbsAssistant v. `r packageVersion("bbsAssistant")` [@bbsAssistant1], 
and tidyverse v. `r packageVersion("tidyverse")` [@tidyverse].
Packages drake v. `r packageVersion("drake")` [@drake], 
bookdown `r packageVersion("bookdown")` [@bookdown1] and
renv v. `r packageVersion("renv")` [@renv] were used to ensure reproducibility.
Code to reproduce all the analyses is archived at [https://github.com/richardjtelford/count.check.ms](https://github.com/richardjtelford/count.check.ms) and [zenodo.org](zenodo.org).
All issues identified in the Neotoma database were reported to the maintainers.

# Results

## Prevalence of singletons

The North American breeding bird survey is a very large, annual monitoring programme to track the status of North American bird populations [@bbsAssistant2].
All birds seen in point counts along ~40 km survey routes at the height of the breeding season are counted.
The very large number of routes, over `r format(floor(bird_singleton_summary$n_routes[1]/1000) * 1000, big.mark = ",")`, typically with large numbers of individuals, allows the ideas developed above to be tested: do most routes have singletons and is the greatest common divisor one, and identify circumstances for which they fail.
This can be tested for the observed data, and by simulating smaller counts and lower richness.

The vast majority of the bird routes have singletons at the species level.
Only `r bird_summ$n_without_singletons` routes do not have singletons, of these only `r bird_summ$n_GCD_gt1 %>% nrow() %>% words()`, all with very low richness, have a greatest common divisor greater than one.

To explore the effect of taxonomic richness on the prevalence of singletons, I aggregate the data to progressively lower taxonomic resolutions.
As richness declines, the proportion of routes having singletons declines (Table \@ref(tab:bird-table)), reaching a moderate proportion at the order level, where most of the counts lacking singletons have fewer than five taxa.
The vast majority of counts have a greatest common divisor of one, except when taxonomic richness is low.

```{r bird-table}
options(digits = 4)
bird_singleton_summary %>%
  select(`Taxonomic level` = taxonomic_level, `Percent with singletons` = p_singletons, `Percent GCD = 1` = GCD1, `Median richness` = median_richness) %>%
  knitr::kable(digits = c(0, 6, 6, 3), caption = "Percent of bird counts from the North American breeding bird survey with singletons and of greatest common divisor (GCD) of one, at different taxonomic levels to mimic the effect of taxonomic richness.")
options(digits = 2)
```

To test the sensitivity of the prevalence of routes without singletons to the count sum, I resample the routes with 400 or more individuals to smaller count sums using a multivariate hypergeometric distribution.
At the order level, the proportion of counts with singletons increases steeply until it reaches a maximum at about 200, thereafter is declines slowly (Fig. \@ref(fig:bird-recount)).
The prevalence of counts with singletons is low for small counts because it is possible that only the common taxa are counted; with larger counts, the chance of counting a single individual of a rare taxon increases.
With even larger counts the prevalence of routes with singletons decreases due to saturation as singletons become doubletons and few new taxa are found.

```{r bird-recount, fig.cap="Effect of count sum on the proportion of bird counts from the North American breeding bird survey without singleton at the order level. Results are the mean of ten trials, error bars are two standard errors.", fig.width = 80/25.4, fig.height = 70/25.4}
readd(bird_recount_singleton_plot)
```

The ` nrow(owens_summ1)` diatom counts from Owen's Lake span a wide range of count sums (` min(owens_summ1$count_sum)` -- ` max(owens_summ1$count_sum)`) because, if concentrations were low, the requirement that at least 300 valves were counted was replaced by a minimum area of slide to count [@Platt1997].
Discounting the ` sum(owens_summ1$n_taxa == 1)` mono-specific assemblages (all with very low count sums ` max(owens_summ1$count_sum[owens_summ1$n_taxa == 1]) %>% insist(. <= 10) %>% invisible()`), ` owens_summ$p_singletons`% of assemblages have singletons with a median number of singletons per assemblage of ` owens_summ$median_sing %>% words()`.
Assemblages with count sums above 50 have a higher chance of having singletons (` owens_summ$no_single_gt_50`% vs ` owens_summ$no_single50_`%).
` owens_summ$gcd2`% of assemblages have a greatest common divisor of one, the remainder have a greatest common divisor of ` owens_summ$gcd_max %>% words()`.
Assemblages with a greatest common divisor above one all have low taxonomic richness` owens_summ1 %>% filter(gcd > 1) %>% pull(n_taxa) %>% max() %>% insist(. <= 6) %>% invisible()`.

The `r testate_summ$n_assemblages` assemblages in the North American testate amoeba training set with available count data [@AMESBURY2018] have count sums between `r testate_summ$count_min` and `r testate_summ$count_max` tests (median = `r testate_summ$count_median`), and taxonomic richness ranges between `r testate_summ$tax_min` and `r testate_summ$tax_max` (median = `r testate_summ$tax_median`) taxa.
`r testate_summ$p_singletons`% of the assemblages have at least one singleton (median `r testate_summ$median_singletons`) and `r testate_summ$p_gcd1`% of assemblages have a greatest common divisor of one.
Of the `r testate_summ$n_gcd1 %>% words()` assemblages with a greatest common divisor above one, `r testate_summ$n_gcd2_low_div %>% words()` have fewer than five taxa; `r testate_summ$n_gcd2 %>% words()` have a greatest common divisor of two, and the other `r testate_summ$n_gcd3 %>% words()` has a greatest common divisor of three.

```{r multitaxa}
options(digits = 4)
readd(count_summary) %>% 
  knitr::kable(caption = "Properties of different data sets")
options(digits = 2)
```

The pollen data from the Neotoma database included over `r pollen_summ$n_assemblages` assemblages in `r pollen_summ$n_datasets %>% format()` datasets. 
Conforming to expectations, nearly all assemblages have singletons (`r pollen_summ$p_singletons`%) and a greatest common divisor of one (`r pollen_summ$p_gcd1`%). 
The assemblages without singletons or with a greatest common divisor above one are not randomly distributed among the datasets.
Half of the assemblages without singletons are in `r pollen_summ$half_no_sing` datasets.
`r Words(pollen_summ$half_gcd)` datasets contain half of the `r pollen_summ$n_gcd1` assemblages with a greatest common divisor above one and only `r pollen_summ$dataset_gcd1`% of the datasets have any such assemblages.
Some Neotoma datasets are with a high proportion of assemblages without singletons or a with a greatest common divisor above one are discussed in section \@ref(unexpected-count-data) of this paper.

## Estimating the percent sum

The Owen's Lake dataset [@Platt1997] includes percentages for each taxa, allowing the count sums estimated from the percent data to be verified. Excluding mono-specific assemblages, for which no meaningful estimate of the count sum is possible, the minimum percent method and the direct search method correctly estimate the count sum for, respectively, ` owens_est_summ$minpc`% and ` owens_est_summ$direct`% of the assemblages. 
The few assemblages that fail the direct search method are species poor and have low count sums.

The `r last_chance_summ$nsamples` chironomid head capsule assemblage dataset from Last Chance Lake [@Axford2017] also includes both the count sums and the percentage of each taxa.
The percentages are given to two decimal places: to account for rounding errors, the countSum package uses the smallest and largest values consistent with the reported percentage.
With the minimum percent method the estimated count sums are, within error, either identical to the reported count sum or twice as much (Fig. \@ref(fig:last-chance-plot)).
With the direct search method, the estimated count sums are all exactly twice the reported count sum.
The factor of two difference is because all the counts include half head capsules (but only sometimes is the rarest taxon represented by a half head capsule).
Reporting half microfossils is common for several microfossil groups, including chironomids and pollen (especially for bisaccate conifers); occasionally other fractions are reported.
Half counts make the estimated count too high by a factor of two, so do not risk incorrectly flagging count sums as being too small.
The direct search method is more precise than the minimum percent method as the rounding error is relatively smaller on the larger percent values the method uses.

```{r last-chance-plot, fig.cap = "Estimated and reported chironomid count sums by the minumum percent method (solid symbols with error bars) and the direct search method (red crosses) from Last Chance Lake [@Axford2017]. Lines show the 1:1 (solid) and 2:1 (dashed) relationships.", fig.width = 80/25.4, fig.height = 80/25.4}
readd(last_chance_plot)
```

## Unexpected count data

This section explores some of the pollen datasets in the Neotoma database that have a high proportion of assemblages without singletons or with a greatest common divisor greater than one.

Some of these datasets are older datasets which may, despite the available metadata, have been back-transformed from percent data or digitised with the loss of rare taxa and precision.
For example, all but `r hershop_pollen_summ$has_singleton %>% word()` of the `r hershop_pollen_summ$n` assemblages in the archived pollen data from Hershop Bog, Texas have a greatest common divisor of two or more (the exception has a single singleton and is otherwise divisible by two). 
The original data [@Patty1968] were counted to exactly 200 pollen grains and fern spores. 
It seems likely that the data were converted to percent, rounded and then back-transformed to counts.
For most purposes, such back-transformed or digitised data are perfectly adequate.

Some datasets with assemblages without singletons have few taxa.
For example, the `r wildhorse$n_assemblage` assemblage dataset from Wildhorse Lake [@Mehringer1985] includes only `r wildhorse$n_taxa_dataset` taxa (`r mean(pollen_summ2$n_taxa_dataset > wildhorse$n_taxa_dataset) * 100` % of Neotoma pollen datasets have more than this), and relatively large count sums (median `r wildhorse$median_count`). 
While `r wildhorse$no_singletons * 100`% of assemblages in this dataset lack singletons, even more assemblage (`r wildhorse$no_zeros * 100`%) lack taxa with zero counts. 
This is in contrast to most (`r pollen_summ$has_zeros * 100`%) Neotoma pollen assemblages which have zero counts.  
These patterns could arise either when there is a high sampling effort in a low diversity site, or when the archived data do not include the entire pollen assemblage, instead focusing on the subset of abundant taxa of most interest to the authors.

For some other datasets, the reason that assemblages lack singletons or have a greatest common divisor above one is less clear. 
These are treated anonymously.

`r Words(pollen1$n_gcd2)` of the `r pollen1$n_sample` assemblages in dataset pollen1 lack singletons and have a greatest common divisor of two.
These assemblages have between `r pollen1$min_taxa` and `r pollen1$max_taxa` taxa.
The probability of all _n_ counts being divisible by _k_ is $k^{-n}$, assuming the distribution of counts is locally uniform.
That is, for _k_ = 2, odd counts are as common as even counts.
For an assemblage with `r pollen1$min_taxa` or `r pollen1$max_taxa` taxa, the probability of having all counts divisible by two by chance are 10^`r ceiling(- pollen1$min_taxa * log10(2))`^ and 10^`r ceiling(- pollen1$max_taxa * log10(2))`^ respectively.
These probabilities are not so small that it is unrealistic to expect to find assemblages like these in the Neotoma database, but to find so many in one dataset is unexpected.
Furthermore, the assumption made here is optimistic, in reality for _k_ = 2, more odd counts are expected ones are expected than even ones in most datasets, as typical rank abundance curves have more singletons than doubletons and so on, which makes finding assemblages with a greatest common divisor above one even more unlikely.

Dataset pollen2 has `r pollen2$n_sample` assemblages, of which `r words(pollen2$n_gcd2)` have a greatest common divisor of two and a further `r words(pollen2$n_gcd3)` with a greatest common divisor of three, one of which has `r pollen2$max_taxa3` taxa.
The probability of having all counts divisible by three in such a taxonomically rich assemblage is 10^`r ceiling(- pollen2$max_taxa3 * log10(3))`.
In addition to the assemblages with a greatest common divisor above one, this dataset includes several assemblages that have a higher proportion of counts divisible by two or three than is typical in assemblages from Neotoma (Fig. \@ref(fig:multiples-plot)).
It is unclear what processes could generate such counts.

```{r multiples-plot, warning = FALSE, fig.height = 100/24.5, fig.width=180/24.5, fig.cap = "A) Distribution of the percent of counts that are divisible by _k_ in each pollen assemblage with at least ten taxa. The vertical dashed line represents the 99.5% percentile. B) Percent of count in each assemblage in dataset pollen2 that are divisible by _k_ against depth. The dashed lines are the quantiles shown in part A. ", message = FALSE}
readd(multiples_plot)
```

In dataset pollen3, the greatest common divisor is three for all assemblages: all `r pollen3$n_count` counts in the dataset are divisible by three (Table \@ref(tab:unexpected-table)). 
For a dataset this large, the probability of having all counts divisible by three by chance is 3^{-`r pollen3$n_count`}^ or 10^`r ceiling(- pollen3$n_count * log10(3))`^.
Given this very small probability, it seems reasonable to conclude that the archived data are not the original counts.

Many assemblages in datasets pollen4 and pollen5 lack singletons and have a greatest common divisor greater than one (Table \@ref(tab:unexpected-table)).
For most of both datasets, which come from the same paper and match the published pollen stratigraphies, assemblages with and without singletons are arranged alternately.
Assemblages without singletons are on the 10 cm intervals, and those with on the intervening 5 cm intervals. 
Closer inspection reveals that the assemblages between two assemblages without singletons have counts that are the mean of their neighbours.
On request, the authors sent the original count data. 
Comparison of these with the archived data showed that the archived assemblages without singletons had had their counts multiplied by two.
The intervening assemblages did not exist in the original data. 
In the published and archived data, they had been interpolated from their neighbours (Fig. \@ref(fig:interpolated-plot)).
The effect of these manipulations is to increase the resolution of the pollen stratigraphy.

```{r unexpected-table}
readd(unexpected_table) %>% knitr::kable(caption = "The number of assemblages, median number of taxa and count sum, percent of assemblages with a greatest common divisor (GCD) of one, in five anonymised pollen datasets.")
```

```{r interpolated-plot, warning = FALSE, fig.height = 100/24.5, fig.width=180/24.5, fig.cap="A) Plot of selected taxa by depth from dataset pollen2. Counts with blue triangles are from assemblages with a greatest common divisor of one. These appear to have been linearly interpolated between the adjacent counts from assemblages with a greatest common divisor of two (red circles). B) Percent of taxa in each assemblage where the count is the mean of the counts in the adjacent assemblages, excluding runs of zero counts. Most of the assemblages with a greatest common divisor of one are exactly linearly interpolated between adjacent assemblages. The main exception is the assemblage at 210 cm which has a single singleton and otherwise has a greatest common divisor of two, other exceptions have a few extra singletons not predicted from the adjacent assembages.", message = FALSE}
readd(pollen4_plot)
```

## Unexpected percent data
```{r line169}

```

Dataset diatom1 includes `r diatom1_summ$nsamples` diatom assemblages with a median richness of `r diatom1_summ$median_taxa` taxa.
The associated paper reports that at least 400 valves were counted from each assemblage, which means that singletons should have a relative abundance of 0.25% or less.
However, the rarest taxon in two assemblages have a relative abundance of `r diatom1_summ$maxpc`%, and the relative abundance of all other taxa in these assemblages are, within rounding error, integer multiples of this`r diatom1_summ$prop_integer %>% insist(. == 1) %>% invisible()`.
If the count sums actually are at least 400, this would imply that each species count in these assemblages is an integer multiple of at least `r diatom1_summ$integer_mult`.
This seems unlikely.
Alternatively, the actual count sum for these assemblages could be as low as `r diatom1_summ$low_n` valves.
In total, `r diatom1_summ$n_low_n %>% words()` assemblages appear to have a count sum below 400 valves. 
The results section of the paper acknowledges that counts were below the standard 400 valves for part of the Late Glacial.
All the assemblages identified with possible low count sums are in this part of the stratigraphy, demonstrating the utility of the method.

```{r line178}

```

The archived data for diatom1 include `r diatom1_summ$n_taxa` taxa that have a maximum abundance of at least 1.25%; this is about half of the 109 taxa reported in the paper.
This pruning of rare taxa will increase the risk of counts without singletons, which would cause the minimum percent method to fail, but will have little impact on the direct search method.

Dataset diatom2 has `r diatom2_summ$nsamples` assemblages with a median richness of `r diatom2_summ$median_taxa` taxa.
The associated paper reports that the diatom counts included 400–500 valves per assemblage except for four diatom-poor assemblages with at least 100 valves.
The direct search method estimates that `r diatom2_summ$n400 %>% words()` assemblages have a count sum of less than 400, and `r diatom2_summ$n100 %>% words()` have less than 100.
The direct search and minimum percent methods agree, within rounding error, `r diatom2_summ$meth_agree %>% insist(isTRUE(.)) %>% invisible()`on the estimated count sums, implying that either all assemblages have singletons and count sums are low in some assemblages, or that the greatest common divisor of the raw counts is greater than one.
This means that for the assemblage where the minimum percent is `r diatom2_summ$maxpc`, which has `r diatom2_summ$rich_maxpc %>% words()` taxa, all counts would need to be multiples of `r diatom2_summ$integer_mult` for the count sum to be at least 100.

Dataset chironomid1 has `r chironomid1_summ$nsamples` assemblages.
Although the associated paper reports that assemblages with count sums below 50 were discarded, the direct search method estimates that `r chironomid1_summ$n_below50 %>% words` assemblages have count sums between `r chironomid1_summ$min_below50` and `r chironomid1_summ$max_below50`.
These assemblages are diverse (richness between `r chironomid1_summ$richness %>% paste(collapse = " and ")` taxa), so it is unlikely that the true counts are double that estimated here and the greatest common divisor is two.

Palaeoceanographic dataset marine1 has `r marine1_summ$nsamples` assemblages with a median taxonomic richness of `r marine1_summ$median_taxa` taxa.
The associated paper reports that assemblages with count sums below 100 were omitted.
The direct search method estimates that `r marine1_summ$n100 %>% words()` assemblages have count sums below 100 (this excludes `r marine1_summ$n_one_taxa %>% words()` mono-specific assemblage), with count sums as low as `r marine1_summ$est_n_min`.
With this dataset, the direct search and the minimum percent method give identical results for some assemblages but highly divergent estimates for others, with the direct search method sometimes giving estimates of several thousand, in one case over two orders of magnitude higher than the minimum percent method.
Some of the divergence appears to be because some, mainly low diversity, assemblages lack singletons.
The most extreme divergences appear to be data entry errors with incorrect rounding, perhaps during taxonomic revisions of the percent data rather than the raw counts.

# Discussion

Analysis of the breeding bird, Owen's Lake diatom, testate amoeba and pollen datasets has shown that, as expected, the vast majority of community and assemblage counts have singletons, and an even larger proportion have a greatest common divisor of one. 
Therefore, these characteristics are potentially useful for identifying data where the count sum is misreported.

My search through archived microfossil data found several datasets where count sums are, or appear to be, smaller than that reported.
Some of the count sums appear to be an order of magnitude below what was reported.
The non-pollen datasets examined here are a convenience sample of available data that met the inclusion criteria.
As such, this analysis cannot be used to accurately determine the prevalence of datasets with possible undercounts, but it appears that a small but non-negligible fraction of the literature is affected.
It is possible in some cases that the assemblages with low counts were omitted or merged prior to analysis, but the number of assemblages reported was not updated.
Care was taken to identify any such cases by, for example, examining stratigraphic diagrams to see if the low count assemblages were included.
In some cases, the assemblages with low apparent counts can be seen in the published stratigraphic diagrams.

It is also possible that the apparently low count sums are a false positive. 
However, as taxonomically diverse assemblages with a greatest common divisor greater than one are rare and all the datasets with apparently low count sums contained several such assemblages, this is unlikely.

These tests require the assemblage data which are often not archived. 
However, if the count sums are very low (below about 20) it is possible to recognise this from a stratigraphic diagram.

## Challenges

Microfossil percent data are often given to two decimal places. 
This is sufficient precision for the direct search method to have utility with counts sums of several thousand. 
Some data are only given to the nearest percent, which means that neither method has utility with count sums above 100.

The direct search method will fail if the dataset includes taxa calculated with different count sums, for example pollen sums of trees, shrubs and upland herbs and a pollen and spore sum that also includes pteridiophytes. 
It should be possible to identify such cases from the meta-data and knowledge of usual practice with different proxies. 
It will also fail, as shown by dataset marine1, if percentages are incorrectly calculated or rounded.

Some taxonomic groups have microfossils that often come in groups of attached individuals.
For example, each diatom cell has two valves which are the counting unit.
These valves usually separate during processing, but paired valves are often found, and some taxa such as _Aulacoseira_ produce long chains of strongly bound valves.
This might make singletons slightly less likely. 
The results from Owen's Lake suggest that this is not an important problem, as most assemblages have many singletons.

## Consequences

Small undercounts in a few assemblages will have minimal impact on the precision of any palaeoenvironmental reconstruction or other statistics derived from the assemblage.
Substantial undercounts will potentially seriously effect the precision of the results.
Such undercounts might constitute a data handling error, for example, if assemblages with low counts were supposed to be merged or omitted but that step was forgotten.
Substantial and pervasive under-counting could be construed as scientific misconduct due to either negligence or falsification, and action to correct the literature is probably required.

Some papers do not report count sums.
When this important quality metric is omitted, the reader should be able to assume that the standard minimum count sum for the taxonomic group has been used (i.e. 50 for chironomids, several hundred for pollen and diatoms).
The methods in this paper may help estimate what the count sums are.
If the actual count sums are materially below this, then action to correct the literature may be necessary, and if the omission is deliberate it 
potentially constitutes misconduct [@Fanelli2013].

# Conclusions

A small but non-negligible fraction of archived assemblage data includes counts that appear, using methods developed above, to have a count sum below that reported in the original paper. 
Some of the apparent counts are small enough that the uncertainty of the count and derived statistics will be substantially larger than expected.
These results highlight the importance of archiving both the raw microfossil count data and the code required to process the data.

# Acknowledgements{-}

Some data were obtained from the Neotoma Paleoecology Database (http://www.neotomadb.org), and the work of the data contributors and the Neotoma community is gratefully acknowledged. This work was partly supported by Norwegian Research Council project PalaeoDrivers (213607).

```{r}
#set digits to default
options(digits = 7)
```

# References{-}